---
title: "zach_ufos"
author: Zach Bogart (zb2223)
output: html_document
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE, eval = FALSE)
knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE, 
                      cache = TRUE)
```

## Introduction
*Explain why you chose this topic, and the questions you are interested in studying.
List team members and a description of how each contributed to the project.*

We were interested in looking at UFO sightings.

```{r}
library(tidyverse)

ufo <- read_csv("ufo_data.csv")
```


## Clean Data

Let's format `Posted` as a Date
```{r}
# Posted

ufo2$Posted <- as.Date(ufo2$Posted, format = "%m/%d/%y")
```

Let's clean the `Shape` column. 
```{r}
# clean Shape
ufo$Shape <- tolower(ufo$Shape)
ufo$Shape[ufo$Shape == "triangular"] <- "triangle"
ufo$Shape[ufo$Shape == "changed"] <- "changing"

```

I added the `In_USA` column to check if the sighting occured in the US or not
```{r}
# Location

in_usa <- function(value) {
  states <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DC", "DE", "FL", "GA", 
          "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", 
          "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", 
          "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 
          "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY")
  if (value %in% states) {
    return(TRUE)
  }
  else {
    return(FALSE)
  }
}

ufo <- ufo %>%
  mutate(`In_USA` = lapply(ufo$State, in_usa))
```

Here is a quick and dirty pass at cleaning the durations
```{r}
# Duration
ufo2 <- ufo

ufo2$Duration <- tolower(ufo2$Duration)
ufo2$Duration_Raw <- ufo2$Duration

# make all spaces single
ufo2$Duration_Raw <- gsub("[[:space:]]+", " ", ufo2$Duration_Raw)
# Remove approx versions
ufo2$Duration_Raw <- gsub("^approx", "", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("^aprox", "", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("^about", "", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("^abt", "", ufo2$Duration_Raw)
# Remove Symbols from beginning of all strings
ufo2$Duration_Raw <- gsub("^[[:punct:]]*", "", ufo2$Duration_Raw)
# Remove Symbols from end as well
ufo2$Duration_Raw <- gsub("[[:punct:]]*$", "", ufo2$Duration_Raw)
# For ranges of time, use minimum value
ufo2$Duration_Raw <- gsub("([0-9]+)-([0-9]+)", "\\1", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("([0-9]+)[[:space:]]+to[[:space:]]+([0-9]+)", "\\1", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("([0-9]+)[[:space:]]+-[[:space:]]+([0-9]+)", "\\1", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("([0-9]+)[[:space:]]+or[[:space:]]+([0-9]+)", "\\1", ufo2$Duration_Raw)
# change digits
ufo2$Duration_Raw <- gsub("one", "1", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("two", "2", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("three", "3", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("five", "5", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("six", "6", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("ten", "10", ufo2$Duration_Raw)
# word approximations to digits
ufo2$Duration_Raw <- gsub("few", "5", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("sew", "5", ufo2$Duration_Raw)
ufo2$Duration_Raw <- gsub("several", "7", ufo2$Duration_Raw)

# left-over plus issue
ufo2$Duration_Raw <- gsub("\\+", "", ufo2$Duration_Raw)
# Add spaces if needed between number and time
ufo2$Duration_Raw <- gsub("([0-9]+)([a-z]+)", "\\1 \\2", ufo2$Duration_Raw)
# remove any starting spaces
ufo2$Duration_Raw <- gsub("^[[:space:]]+", "", ufo2$Duration_Raw)

# make all spaces single
ufo2$Duration_Raw <- gsub("[[:space:]]+", " ", ufo2$Duration_Raw)
```

```{r}
duration_clean_first_pass <- function(value) {
  if (!is.na(value)) {
    import <- strsplit(value, " ")
    times <- paste(import[[1]], sep = " ")
    # if the first value is just a number
    if (suppressWarnings(!is.na(as.numeric(times[1])) & !is.na(times[2]))) {
      # trim whitespace from items
      times[1] <- trimws(times[1])
      times[2] <- trimws(times[2])
      
      if (times[2] %in% c("second", "seconds", "secs.", "secs", "sec.", "sec", ",seconds")) {
        result <- as.numeric(times[1])
      } else if (times[2] %in% c("minute", "minutes", "min", "min.", "nins", "mins", "minutes.", 
                                 "minutew", "minuets", "minuts", "minutes,", "minutos", "miutes",
                                 "mint", "mintues", "ninutes", "minustes", "mn", "mis", "mini", 
                                 "muntes", "minutue", "minets")) {
        result <- as.numeric(times[1]) * 60
      } else if (times[2] %in% c("hour", "hours", "H", "h", "hrs", "hrs.", "hr", "hpurs")) {
        result <- as.numeric(times[1]) * 3600
      } else if (times[2] %in% c("day", "days")) {
        result <- as.numeric(times[1]) * 86400
      } else if (times[2] %in% c("week", "weeks")) {
        result <- as.numeric(times[1]) * 604800
      } else if (times[2] %in% c("night", "nights")) {
        result <- as.numeric(times[1]) * 28800 # assume "night" is 8 hours
      } else {
        result <- paste(times, collapse = " ")
      }
    } else if (value %in% c("hour", "hours", "hr")) {
        result <- 3600 # if "hour", make 3600 seconds
    } else if (value %in% c("short", "brief", "minute", "minutes", "minets", "less than 1 minute", "less than a minute")) {
        result <- 60 # if "short", make 1 minute
    } else if (value %in% c("seconds", "few seconds", "a few seconds")) {
        result <- 10 # if "seconds", make 10 seconds
    } else if (value %in% c("unknown", "?", "???", "not known", "unkn", "don't know", "unk", 
                            "don't remember", "unsure", "not sure", "uncertain", "unknow", "not given",
                            "unclear")) {
        result <- NA
    } else {
      result <- paste(times, collapse = " ")
    }
  }
  else {
    result <- NA
  }
  return(result[[1]][1])
}

ufo2$Duration_Cleaned <- lapply(ufo2$Duration_Raw, duration_clean_first_pass)
```

To see how we are doing, let's see how many new NAs are introduced
```{r}
ufo2$Duration_Testing <- ufo2$Duration_Cleaned

cat("Before: ", sum(is.na(ufo2$Duration_Testing)))

ufo2$Duration_Testing <- as.numeric(ufo2$Duration_Cleaned)

cat(" After: ", sum(is.na(ufo2$Duration_Testing)))

```

```{r}
duration <- ufo2$Duration_Testing[!is.na(ufo2$Duration_Testing)]

summary(duration)

str(unique(ufo2$Duration_Cleaned))
```


```{r}
cleaned <- read_csv("ufo_clean.csv")
standard <- read_csv("ufo_data.csv")
```










